---
title: "Work_Sample_Analysis_Clustering"
author: "Heena"
date: "April 1, 2018"
output:
  html_document: default
---

```{r}
cal <- read.csv("Calendar.csv", header = TRUE)
ft <- read.csv("FileType.csv", header = TRUE)
task <- read.csv("Task.csv", header = TRUE)
uni <- read.csv("Universe.csv", header = TRUE)
```

```{r}
# packages to be preinstalled with the libraries used in this project.

#install.packages("ggplot2")
library(ggplot2)
#library(ggthemes)

#install.packages("dplyr")
library("dplyr")


library("scales")

library("lubridate")

library("stringr")

#install.packages("factoextra")
library("factoextra")
```

```{r}
summary(cal)
summary(cal$Billable.CE)    # it contains numerical values with values between 0 and 5 and is depended on Actual.CE as given, 
summary(cal$Billable.CE > 5)  # there are 1745 values greater than 5 which shouldn't be there, let's drop those rows
cal <- cal[!cal$Billable.CE > 5,]

sum(cal$Supplier == "Unknown")  # we will remove Unknown at required places

# Also Developer.ID and eMail are same, so one can be dropped

##

summary(ft) 
summary(ft$Actual.CE) # it contains outliers
summary(ft$Actual.CE[!ft$Actual.CE %in% boxplot.stats(ft$Actual.CE)$out])  # without outliers

# eMail can be used instead of Developer.ID over here

##

summary(task)

##

summary(uni)
```



***
***

## a # Number of Vendors used in each country

```{r}
str(cal$Supplier)       # it has 45 known vendors and the rest suppliers fall under unknown
summary(cal$Supplier == 'Unknown')  # there are 134055 suppliers which do not fall in given vendors

# as we don't need unknown values over here, let's drop them

a <- aggregate(Supplier ~ Country, data = cal, length)

a_2 <- a[!a$Country == "Unknown", ] # this drops all the rows containing Unknown supplier
```

```{r}
ggplot(data = a_2, aes(Country, Supplier, fill = Supplier)) + geom_bar(stat = "identity") + 
        xlab("Country") + ylab("Supplier Count") +
        ggtitle("Number of Vendors used in each country") +
        theme(panel.background = element_rect(fill = "white",
        color = "black", size = 0.1),
        axis.text.x = element_text(angle = 45, hjust = 1)) +
        scale_y_continuous(breaks = pretty_breaks(n = 10)) 
```


***
***

## b # the average number of Developers in all the repositories and What is the average number of Repositories in all the Projects

```{r}
# Each repository has a unique Infra ID, so we take InfraId for our calculation as Calendar dataset contains both

# let's look at the structure of InfraId, Developer.ID and Project
str(cal$InfraId)        # it is an integer value,let's see number of unique values
length(unique(cal$InfraId)) # it contains 1589 unique values
str(cal$Developer.ID)   # it contains 4212 different developers
str(cal$Project)        # it contains 171 projects
```

```{r}
# the average number of Developers in all the repositories 

repo_dev <- aggregate(Developer.ID ~ InfraId, data = cal, length)

(sum(repo_dev$Developer.ID)/ length(repo_dev$InfraId))
```


```{r}
# the average number of Repositories in all the Projects

proj_repo <- aggregate(InfraId ~ Project, data = cal, length)

(sum(proj_repo$InfraId)/ length(proj_repo$Project))
```


***
***

# c # the Top 3 Suppliers in terms of Billable Coding Effort (CE) Delivered per developer

```{r}
# let's look at the structure of the relevant columns

str(cal$Developer.ID)   # it contains 4212 different developers

str(cal$Billable.CE)    # it contains numerical values with values between 0 and 5 and is depended on Actual.CE as given, 

str(cal$Supplier)       # it has 45 known vendors and the rest suppliers fall under unknown
summary(cal$Supplier == 'Unknown')  # there are 134055 suppliers which do not fall in given vendors

# as we don't need unknown values over here, let's drop them

c <- data.frame(cal$Developer.ID, cal$Billable.CE, cal$Supplier)

c_2 <- c[!c$cal.Supplier == "Unknown", ]  # this drops all the rows containing Unknown supplier
```


```{r}
sup_bill <- aggregate(cal.Billable.CE ~ cal.Supplier + cal.Developer.ID, data = c_2, sum)

# let's arrange sup_bill according to Billabe_CE

# library(dplyr)
sup_bill_10 <- head(arrange(sup_bill, desc(cal.Billable.CE)), n = 10)  # We get top 6 suppliers according to their Billable.CE

# Supplier       Developer.ID      Billable.CE
# Vendor-003	  Developer 77149	    284.38	
# Vendor-027    Developer 78544	    239.94	
# Vendor-002	  Developer 77935	    238.85	
```


```{r}
# library(scales)

ggplot(sup_bill_10,aes(x = cal.Supplier, y = cal.Billable.CE, fill = cal.Developer.ID, group = 1)) +
  geom_line(stat = "identity", aes(colour = cal.Developer.ID), size = 3) +
  scale_y_continuous(breaks = pretty_breaks(n = 10)) +
  labs(title = "Top 10 Suppliers in terms of Billable Coding Effort (CE) Delivered per developer") +
  xlab("Supplier") + ylab("Billable.CE") +
  geom_vline(xintercept = 1, linetype = "dashed") +
  annotate("text", label="238.85", colour="black", x = 1.5, y = 237) +
  geom_vline(xintercept = 2, linetype = "dashed") +
  annotate("text", label="284.38", colour="black", x = 2.5, y = 286) +
  geom_vline(xintercept = 5, linetype = "dashed") +
  annotate("text", label="239.94", colour="black", x = 5.5, y = 240) +
  theme(panel.background = element_rect(fill = "white",
        color = "black", size = 0.2),
        axis.text.x = element_text(angle = 45, hjust = 1))
```


***
***

## d # the Project with a maximum number of Billable CE hours per day delivered per developer

```{r}
# let's look at the structure of the relevant columns

str(cal$Project)  # it contains 171 different Projects

proj_bill <- aggregate(Billable.CE ~ Project + Developer.ID, data = cal, sum)

max(proj_bill$Billable.CE)

#    Project     Developer.ID     Billable.CE
#  Project 155  Developer 79398      340.00

proj_bill_10 <- head(arrange(proj_bill, desc(Billable.CE)), n = 10) 
```

```{r}
library(scales)

ggplot(proj_bill_10,aes(x = Project, y = Billable.CE, fill = Developer.ID, group = 1)) +
  geom_line(stat = "identity", aes(colour = Developer.ID), size = 3) +
  scale_y_continuous(breaks = pretty_breaks(n = 10)) +
  labs(title = "Top 10 Projects in terms of Billable Coding Effort (CE) Delivered per developer") +
  xlab("Projects") + ylab("Billable.CE") +
  geom_vline(xintercept = 3, linetype = "dashed") +
  annotate("text", label="340.00", colour="black", x = 3.5, y = 341.5) +
  geom_vline(xintercept = 4, linetype = "dashed") +
  annotate("text", label="327.96", colour="black", x = 4.5, y = 327.96) +
  geom_vline(xintercept = 4, linetype = "dashed") +
  annotate("text", label="298.32", colour="black", x = 4.5, y = 298.32) +
  theme(panel.background = element_rect(fill = "white",
        color = "black", size = 0.2),
        axis.text.x = element_text(angle = 45, hjust = 1))
```


***
***

## e # the difference on the total Actual CE delivered between July and August in Country 17

```{r}
# let's look at the values

str(ft$Actual.CE) 
# it contains outliers which should be removed

str(ft$Actual.CE[!ft$Actual.CE %in% boxplot.stats(ft$Actual.CE)$out])
# summary without outliers
```


```{r}
country_17 <- data.frame(ft$Actual.CE[ft$Country == 'Country 17'], as.factor(month(dmy(ft$Date)))[ft$Country == 'Country 17'], ft$Country[ft$Country == 'Country 17'])
```

```{r}
# let's remove outliers in Actual.CE
country_17 <- country_17[!country_17$ft.Actual.CE.ft.Country.....Country.17.. %in% boxplot.stats(country_17$ft.Actual.CE.ft.Country.....Country.17..)$out,]
```


```{r}
country_17_7 <- sum(country_17$ft.Actual.CE.ft.Country.....Country.17..[country_17$as.factor.month.dmy.ft.Date....ft.Country.....Country.17.. == 7])
country_17_8 <- sum(country_17$ft.Actual.CE.ft.Country.....Country.17..[country_17$as.factor.month.dmy.ft.Date....ft.Country.....Country.17.. == 8])
```

```{r}
country_17_diff <- (country_17_8 - country_17_7)
country_17_diff     # difference in total Actual CE between July and August in Country 17 is 1275.94
```

```{r}
diff <- c(country_17_8, country_17_7)

barplot(diff,
  main = "Total Actual CE delivered in July and August in Country 17",
  ylim = c(0,8500),
  #horiz = TRUE,
  las = 1,
  names.arg = c("July", "August"),
  col = c("red4","red3"),
  ylab = "Total Actual.CE in Country 17") 

text(0.8, 8457, "8254.12")
text(1.9, 7400, "6978.18")
```

***
***

## f # the total number of developers in each one of the last 3 months for Vendor-002

```{r}
# library("lubridate")
f <- data.frame(cal$Developer.ID, cal$Date, cal$Supplier)
f$only_month <- as.factor(month(dmy(f$cal.Date))) # convert month values to factor from numeric
# OR  (above step can also be done like below)
#f$only_month <- month(as.POSIXlt(f$cal.Date, format = "%d/%m/%Y"))
#f$only_month <- as.factor(f$only_month) 

str(f)  # we can see the new structure of Date in only_month
```

```{r}
ven_002 <- f[f$cal.Supplier == "Vendor-002",]   # we only need Vendor-002 supplier here, so extract that
```


```{r}
# total number of developers in each one of the last 3 months for Vendor-002
total_dev_7 <- length(unique(ven_002$cal.Developer.ID[ven_002$only_month == 7]))  # 368
total_dev_8 <- length(unique(ven_002$cal.Developer.ID[ven_002$only_month == 8]))  # 336
total_dev_9 <- length(unique(ven_002$cal.Developer.ID[ven_002$only_month == 9]))  # 226
```


```{r}
total_dev <- c(total_dev_7, total_dev_8, total_dev_9)

barplot(total_dev,
  main = "No. of developers in last 3 months for Vendor-002",
  ylim = c(0,400),
  #horiz = TRUE,
  las = 1,
  names.arg = c("July", "August", "September"),
  col = c("red4","red3","red2"),
  ylab = "Number of Developers") 

text(0.8, 380, "368")
text(1.9, 350, "336")
text(3.0, 240, "226")

```


***
***

## g # Percentage of the total projects working with Java Files

```{r}
project_java <- ft$Project[ft$FileType == 'java']
prop <- length(project_java)/length(ft$Project)
prop*100    # 25.3 % of total projects worked with Java Files
```

```{r}
ggplot(data = ft, aes(FileType, group = length(FileType))) + geom_histogram(aes(y = ..prop.., fill = factor(..x..)), stat="count") +
  labs(title = "Count of File Type") +
  labs(x = "FileType", y = "Count") +
  scale_y_continuous(labels=scales::percent)
```


***
***

## h # the most common task type in Repository 642

```{r}
# let's look at the structure of the relevant columns

str(uni$Repository)  # it contains 7735 different Repositories

uni$InfraId[uni$Repository == 'Repo 642'] # 41583 InfraId belongs to Repo 642

task_642 <- task$Task.Type[task$InfraId == 41583] # this stores all the Task Types for relevant InfraId

task_642 <- data.frame(table(task_642)) # this gives a task type null with 107 frequency, which should be ignored

task_642 <- task_642[!task_642$task_642 == 'null',] # dropping null

task_642    # most common task is Improvement with frequency of 31
```

```{r}
ggplot(data = task_642, aes(x = task_642, y = Freq, fill = Freq)) + geom_bar(stat="identity") +
  scale_y_continuous(breaks = pretty_breaks(n = 15)) +
  xlab("Task Types In Repo 642") +
  ggtitle("Most common task type in Repository 642") +
  theme(panel.background = element_rect(fill = "white",
        color = "black", size = 0.5),
        axis.text.x = element_text(angle = 45, hjust = 1))
```


***
***
***

# Clustering

```{r}
# store eMail as Developer, as they are same
ft$Developer <- as.factor(gsub("\\@", "", ft$eMail))
```


```{r}
summary(ft)
```


```{r}
#Create a normalization function
# we can normalize the data 

normalize <- function(x) {
return((x - min(x))/(max(x) - min(x)))} # create normalize function
```


```{r}
ft_new <- aggregate(Date ~ InfraId + FileType, data = ft, length) # taking infraid and Filetype and aggregating it under dates for all developers

ft_new$FileType <- as.numeric(factor(ft_new$FileType , levels = c("cc", "cs", "html", "java", "js", "py", "r", "sh", "sql", "xml")))  # change factor to numeric
```

```{r}
ggplot(ft_new, aes(Date, log(InfraId) , color =  FileType)) + geom_point()
# we can observe four classifications for this data
```


```{r}
# normalize 
norm_ft_new <- as.data.frame(sapply(ft_new, normalize))
```


```{r}
# Let's check value for the number of clusters to be taken

# Elbow Method
set.seed(1234)

#library("factoextra")

fviz_nbclust(norm_ft_new, kmeans, method = "wss")
```

```{r}
# silhouette Method
fviz_nbclust(norm_ft_new, kmeans, method = "silhouette")

# We see both methods gives 4 clusters
```



# kmeans

```{r}
# kmeans
cluster_ft <- kmeans(norm_ft_new, centers = 4, nstart = 25) # with 4 clusters

print(cluster_ft) # to extract the results using 4 clusters

str(cluster_ft)
# it outputs a cluster of size 1033, 669, 542 and 1244

fviz_cluster(cluster_ft, data = norm_ft_new)  # to visualize the results
```

```{r}
# Extract the clusters and add to our initial data to do some descriptive statistics at the cluster level:

ft_new %>%
  mutate(Cluster = cluster_ft$cluster) %>%
  group_by(Cluster) %>%
  summarise_all("mean")
```


***
***
***

